Cloud Supercomputing

### 2019-04-27 writing/paper/paper.tex
Fix some small issues.

\begin{document}
\maketitlet 
\begin{abstract}
  In the context of simulation of convective proceess for numerical weather prediction, we present an application of the compressible Navier Stokes equations on cloud formation processes.
  While established methods have to use tailored numerics towards the specific simulation, our generic scheme shows competitive results for both classical cfd and hydrostatic scenarios.
  Our scheme exploits an underlying parallelized implementation of the \aderdg{} method with dynamic adaptive mesh refinement. 
  We improve our method by a PDE independent general refinement criterion, based on the local total variation of the numerical solution.
  All together our method can be seen as a perfect candidate for large scale cloud simulation runs on current and future super computers.
\begin{itemize}
\item Implementation of compressible Navier Stokes with a high-order \aderdg{}-method and a low order finite volume \muscl{}-scheme.
\item Implementation in \exahypeengine{}, a scalable \pde{} framework.
\item Outlier and total variation based adaptive mesh refinement that reduced number of grid cells drastically and preserves solution quality.
\item Simulation of challenging flows over hydrostatic background atmosphere.
\item Competitive results for both classical cfd scenarios and hydrostatic scenarios in two and three dimensions.
\end{itemize}

The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{\aderdg{}  \and Navier-Stokes \and Another keyword.}
\end{abstract}
\section{Introduction}
In this paper we address the resolution of convective processes in the context of numerical weather prediction (\textsc{nwp}), on modern super computer systems.
The simulation of these process, with suitable physical models and on large scale clusters is expected to play an important role in future \textsc{nwp}~\cite{bauer2015quiet}.
%% Numerical weather prediction (\textsc{nwp}) contains processes of different spatio-temporal scales.
%% An important part of future \textsc{nwp} models will be the the resolution of convective processes~\cite{bauer2015quiet}.
%% This part requires both suitable physical models and effective computational realisations. 
We focus on the simulation of cloud formation processes with simple benchmark scenarios~\cite{giraldo2008study}.
These relatively small scale effects are well approximated with the compressible Navier-Stokes equations.
The general approach is to use a space-discretized by higher order Galerkin method, as seen for example in~\cite{giraldo2008study,muller2010adaptive,muller2018strong}.
%%They are often space-discretized by higher order Galerkin methods, as seen for example in~\cite{giraldo2008study,muller2010adaptive,muller2018strong}.
Even though all of these implementations use a discretization with a (theoretical) unlimited space-order, they use Runge-Kutta time integrators or semi-implicit methods.
It is well known that Runge-Kutta methods become more ineffecient for very high orders.
Furthermore, they and the implicit methods are more communication-intensive and thus require a larger overhead and are more complicated to scale.

We follow a different approach in this paper and use the \aderdg{} method of~\cite{dumbser2008unified} to simulate various computational fluid dynamics (\textsc{cfd}) scenarios with a focus on flows with stratified flows.
\todo{LK: Ist hier stratified flows korrekt?}
\todo{MB: viscosity ist unmotiviert ist.
    Lieber noch so 10--15 Zeilen drüber schreiben.}
Due to the viscous components of the Navier-Stokes equations, it is not straightforward to apply the \dg{} method.
For high-order \dg{} methods, viscosity not only models the physical reality of the problem but also helps us by smoothing out oscillations and discontinuities, thus stabilizing the simulation.
To include viscosity, we use the numerical flux for the compressible Navier-Stokes equations of \citeauthor{gassner2008discontinuous}~\cite{gassner2008discontinuous}.
This flux has already been applied to the \aderdg{} method in~\cite{dumbser2010arbitrary}.
In contrast to this paper, we focus on the simulation of complex flows with a gravitational source term and a realistic background atmosphere.
Additionally, we use adaptive mesh refinement (\textsc{amr}) to increase the spatial resolution in areas of interest.
This has been shown to work well for the simulation of cloud dynamics~\cite{muller2010adaptive}.
%\todo{Einordnung, Kontext zu giraldo, etc$\ldots$}
%Insert references to Giraldo SE/DG bubbles~\cite{giraldo2008study}, 3D bubbles~\cite{giraldo2013implicit,kelly2012continuous}, bubble amr (müller~\cite{muller2010adaptive}).

@@

We make the following contributions\todo{Feedback von MB: Besser hervorheben, warum diese Sachen interessant/neu sind.
Dumbser macht in seinen NS-DG papern keine komplexen Source terme, keine hydrostatic flows, etc.
}:
\begin{itemize}%
\item We extend the \exahypeengine{} to allow viscous terms following the aforementioned numerical scheme.
\item We use this to provide an implementation of the compressible Navier-Stokes equation.
  In addition, we tailor the equation set to stratified flows with gravitational source term.
  We want to emphasize here that we use a standard formulation of the Navier-Stokes equations as seen in the field of computational fluid mechanics and only use small modifcations of the governing equations, in contrast to a equation set that is tailored exactly to the application area.
\item We present a general \textsc{amr}-criterion that is based on the detection of outlier cells w.r.t.\ their total variation.
  Furthermore, we show how to utilize this criterion for stratified flows.
\item We evaluate our implementation with standard \textsc{cfd} scenarios and atmospheric flows and inspect the effectiveness of our proposed \amr{}-criterion.
  We thus inspect, whether our proposed general implementation can achieve results that are competitive with the state-of-the-art models that rely on heavily specified equations and numerics.
\end{itemize}

@@

The user can state dynamic \amr{} rules by supplying custom criteria that are evaluated point-wise.
We use an element local error estimate based on the total variation of the numerical solution. 
Let $\bm{f}(\bm{x}): \mathbb{R}^{N_\text{vars}} \to \mathbb{R}$ be a sufficiently smooth function that maps the discrete solution at a point $\bm{x}$ to an arbitrary indicator variable.
The total variation (\textsc{tv}) of this function is defined by
\begin{equation}
  \label{eq:tv}
  \tv \left[ f(\bm{x}) \right] =
  \Vert
\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert }
\Vert_1
\end{equation}
for each cell.
The operator $\Vert \cdot \Vert_1$ denotes the discrete $l_1$ norm in this equation.
We compute the integral efficiently with Gaussian quadrature over the collocated quadrature points.
\todo{LK: If using limiting, add sentence about tv computation for fv}
\newcommand{\mean}{\mu}%
\newcommand{\std}{\sigma}%
\newcommand{\variance}{\std^2}%
\newcommand{\Trefine}{T_\text{refine}}%
\newcommand{\Tdelete}{T_\text{erase}}%
How can we decide whether a cell is important or not?
To resolve this conundrum, we compute the mean and the population standard deviation of the total variation of all cells.
It is important to note that we use the method of~\cite{chan1982updating} to compute the modes in a parallel and numerical stable manner.
A cell is then considered to contain significant information if its deviates from the mean more than a given threshold.
This criterion can be described formally by
\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv(\Q) \geq \mu + \Trefine \sigma, \\
    \text{erase} & \text{if } \tv(\Q) < \mu + \Tdelete \sigma, \\
    \text{keep} & \text{otherwise}.
    \end{cases}
\end{equation}
The parameters $\Trefine > \Tdelete$ can be chosen freely.
Chebyshev's inequality
\begin{equation}
  \label{eq:chebychev}
  \mathbb{P}(\vert X - \mu \vert \geq c \sigma) \leq \frac{1}{c^2},
\end{equation}
with probability $\mathbb{P}$ guarantees that we neither mark all cells for refinement nor for erasement.
\todo{LR: Was gibt es an compressible NS mit dynamic AMR, wie unterscheiden wir uns im Refinement Kriterium?}
This inequality holds for arbitrary distributions under the weak assumption that they have a finite mean $\mu$ and a finite standard deviation $\sigma$~\cite{wasserman2004all}.
Note that subcells are coarsened only if all subcells belonging to the coarse cell are marked for erasion.

We use a \muscl{} based finite volume limiting as proposed in~\cite{dumbser2016simple}.
Briefly, we first compute a candidate solution with our \aderdg{}-scheme and check whether it is valid.
A solution is considered to be invalid if it is a non-finite floating point number or if it is unphysical, i.e.\ if the pressure or density are not positive.
\todo{LK: Mention dmp if using, mention tv-indicator for limiting if using}
In case of an invalid solution, we recompute the timestep with a our \muscl{} finite volume method.\todo{Source for the fv method}

@@

\subsection*{Acknowledgements}
Lukas Krenz was funded by ChEESE (823844), Leonhard Rannabauer was funded by ExaHyPE (671698).
\todo{LK: Standard sentence is: This research was funded by the European Union’s Horizon 2020 Research and Innovation Programme under theproject ExaHyPE, grant no.  671698 (call FETHPC-1-2014).

Also LRZ?}

>>>>

\begin{document}
\maketitle 
\begin{abstract}
  In the context of simulation of convective proceess for numerical weather prediction, we present an application of the compressible Navier Stokes equations on cloud formation processes.
  While established methods have to use tailored numerics towards the specific simulation, our generic scheme shows competitive results for both classical \textsc{cfd} and stratified scenarios.
  Our scheme exploits an underlying parallelized implementation of the \aderdg{} method with dynamic adaptive mesh refinement. 
  We improve our method by a  \pde{} independent general refinement criterion, based on the local total variation of the numerical solution.
  All together our method can be seen as a perfect candidate for large scale cloud simulation runs on current and future super computers.
% \begin{itemize}
% \item Implementation of compressible Navier Stokes with a high-order \aderdg{}-method and a low order finite volume \muscl{}-scheme.
% \item Implementation in \exahypeengine{}, a scalable \pde{} framework.
% \item Outlier and total variation based adaptive mesh refinement that reduced number of grid cells drastically and preserves solution quality.
% \item Simulation of challenging flows over hydrostatic background atmosphere.
% \item Competitive results for both classical cfd scenarios and hydrostatic scenarios in two and three dimensions.
% \end{itemize}

%The abstract should briefly summarize the contents of the paper in
%150--250 words.

\keywords{\aderdg{}  \and Navier-Stokes \and Another keyword.}
\end{abstract}
\section{Introduction}
In this paper we address the resolution of convective processes in the context of numerical weather prediction (\textsc{nwp}) on modern super computer systems.
The simulation of these process, with suitable physical models and on large scale clusters is expected to play an important role in future \textsc{nwp}~\cite{bauer2015quiet}.
%% Numerical weather prediction (\textsc{nwp}) contains processes of different spatio-temporal scales.
%% An important part of future \textsc{nwp} models will be the the resolution of convective processes~\cite{bauer2015quiet}.
%% This part requires both suitable physical models and effective computational realisations. 
We focus on the simulation of cloud formation processes with simple benchmark scenarios~\cite{giraldo2008study}.
These relatively small scale effects are well approximated with the compressible Navier-Stokes equations.
The general approach is to use a space-discretized by higher order Galerkin method, as seen for example in~\cite{giraldo2008study,muller2010adaptive,muller2018strong}.
%%They are often space-discretized by higher order Galerkin methods, as seen for example in~\cite{giraldo2008study,muller2010adaptive,muller2018strong}.
Even though all of these implementations use a discretization with a (theoretical) unlimited space-order, they use Runge-Kutta time integrators or semi-implicit methods.
It is well known that Runge-Kutta methods become more ineffecient for very high orders.
Furthermore, they and the implicit methods are more communication-intensive and thus require a larger overhead and are more complicated to scale.

We follow a different approach in this paper and use the \aderdg{} method of~\cite{dumbser2008unified} to simulate various computational fluid dynamics (\textsc{cfd}) scenarios with a focus on flows with stratified flows.
Due to the viscous components of the Navier-Stokes equations, it is not straightforward to apply the \dg{} method.
For high-order \dg{} methods, viscosity not only models the physical reality of the problem but also helps us by smoothing out oscillations and discontinuities, thus stabilizing the simulation.
To include viscosity, we use the numerical flux for the compressible Navier-Stokes equations of \citeauthor{gassner2008discontinuous}~\cite{gassner2008discontinuous}.
This flux has already been applied to the \aderdg{} method in~\cite{dumbser2010arbitrary}.
In contrast to this paper, we focus on the simulation of complex flows with a gravitational source term and a realistic background atmosphere.
Additionally, we use adaptive mesh refinement (\textsc{amr}) to increase the spatial resolution in areas of interest.
This has been shown to work well for the simulation of cloud dynamics~\cite{muller2010adaptive}.
%\todo{Einordnung, Kontext zu giraldo, etc$\ldots$}
%Insert references to Giraldo SE/DG bubbles~\cite{giraldo2008study}, 3D bubbles~\cite{giraldo2013implicit,kelly2012continuous}, bubble amr (müller~\cite{muller2010adaptive}).

@@

We make the following contributions:
\begin{itemize}%
\item We extend the \exahypeengine{} to allow viscous terms following the aforementioned numerical scheme.
\item We use this to provide an implementation of the compressible Navier-Stokes equation.
  In addition, we tailor the equation set to stratified flows with gravitational source term.
  We want to emphasize here that we use a standard formulation of the Navier-Stokes equations as seen in the field of computational fluid mechanics and only use small modifcations of the governing equations, in contrast to a equation set that is tailored exactly to the application area.
\item We present a general \textsc{amr}-criterion that is based on the detection of outlier cells w.r.t.\ their total variation.
  Furthermore, we show how to utilize this criterion for stratified flows.
\item We evaluate our implementation with standard \textsc{cfd} scenarios and atmospheric flows and inspect the effectiveness of our proposed \amr{}-criterion.
  We thus inspect, whether our proposed general implementation can achieve results that are competitive with the state-of-the-art models that rely on heavily specified equations and numerics.
\end{itemize}

@@

The user can state dynamic \amr{} rules by supplying custom criteria that are evaluated point-wise.
We use an element local error estimate based on the total variation of the numerical solution. 
\todo{Mention intuition, i.e.\ tvd stuff}
Let $\bm{f}(\bm{x}): \mathbb{R}^{N_\text{vars}} \to \mathbb{R}$ be a sufficiently smooth function that maps the discrete solution at a point $\bm{x}$ to an arbitrary indicator variable.
The total variation (\textsc{tv}) of this function is defined by
\begin{equation}
  \label{eq:tv}
  \tv \left[ f(\bm{x}) \right] =
  \Vert
\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert }
\Vert_1
\end{equation}
for each cell.
The operator $\Vert \cdot \Vert_1$ denotes the discrete $l_1$ norm in this equation.
We compute the integral efficiently with Gaussian quadrature over the collocated quadrature points.
%\todo{LK: If using limiting, add sentence about tv computation for fv}
\newcommand{\mean}{\mu}%
\newcommand{\std}{\sigma}%
\newcommand{\variance}{\std^2}%
\newcommand{\Trefine}{T_\text{refine}}%
\newcommand{\Tdelete}{T_\text{erase}}%
How can we decide whether a cell is important or not?
To resolve this conundrum, we compute the mean and the population standard deviation of the total variation of all cells.
It is important to note that we use the method of~\cite{chan1982updating} to compute the modes in a parallel and numerical stable manner.
A cell is then considered to contain significant information if its deviates from the mean more than a given threshold.
This criterion can be described formally by
\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv(\Q) \geq \mu + \Trefine \sigma, \\
    \text{erase} & \text{if } \tv(\Q) < \mu + \Tdelete \sigma, \\
    \text{keep} & \text{otherwise}.
    \end{cases}
\end{equation}
The parameters $\Trefine > \Tdelete$ can be chosen freely.
Chebyshev's inequality
\begin{equation}
  \label{eq:chebychev}
  \mathbb{P}(\vert X - \mu \vert \geq c \sigma) \leq \frac{1}{c^2},
\end{equation}
with probability $\mathbb{P}$ guarantees that we neither mark all cells for refinement nor for erasement.
\todo{LR: Was gibt es an compressible NS mit dynamic AMR, wie unterscheiden wir uns im Refinement Kriterium?}
This inequality holds for arbitrary distributions under the weak assumption that they have a finite mean $\mu$ and a finite standard deviation $\sigma$~\cite{wasserman2004all}.
Note that subcells are coarsened only if all subcells belonging to the coarse cell are marked for erasion.

We use a \muscl{} based finite volume limiting as proposed in~\cite{dumbser2016simple}.
Briefly, we first compute a candidate solution with our \aderdg{}-scheme and check whether it is valid.
A solution is considered to be invalid if it is a non-finite floating point number or if it is unphysical, i.e.\ if the pressure or density are not positive.
%\todo{LK: Mention dmp if using, mention tv-indicator for limiting if using}
In case of an invalid solution, we recompute the timestep with a our \muscl{} finite volume method.\todo{Source for the fv method}

@@

\subsection*{Acknowledgements}
Lukas Krenz was funded by ChEESE (823844), Leonhard Rannabauer was funded by ExaHyPE (671698).
\todo{LK: Standard sentence is: This research was funded by the European Union’s Horizon 2020 Research and Innovation Programme under theproject ExaHyPE, grant no.  671698 (call FETHPC-1-2014).
Special thanks goes to Dominic E.\ Charrier who provided help with the implementation in the \exahypeegine{}.
Also LRZ?}
